---
title: "All-Class Embedding Development"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10,fig.height = 8,warning = FALSE,message = FALSE)
```

The goal of this markdown document is to try to figure out a good way of creating a meaningful cross-class embedding. If we can get this to work it will help us to visualize cell-cell signaling in situ.

This is a challenging task historicall because we are using class-wise integration embeddings to look at each individual class. And we cannot integrate all the data together with default settings because there are so many sources of variance in the data, it creates a big ugly blob that show all the classes smooshed together. No fun.

But what if we force the embedding algoithm to work for us? Might be able to get it to happen, not sure. Let's find out.

#### Set up workspace
```{r}
# Require packages
require(Seurat)
require(NICHES)
require(cowplot)

# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")

# Load color palette
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")
color_pals
```

#### Load the data from Part 1 that was used to calculate NICHES
```{r}
load('downsampled.imputed.Robj')
## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(downsampled$Final2)
```

#### Make smaller for testing

```{r}
Idents(downsampled) <- downsampled$orig.ident
table(Idents(downsampled))
small <- subset(downsampled,cells = WhichCells(downsampled,downsample = 500))
table(Idents(small))
```

The most important variance that we want the embedding algoirthm to take seriously is the information that distingiuishes the 4 classes from each other. That should help us to make pretty cross-class connectivity pictures. Let's try.

```{r}
table(small$class)
Idents(small) <- small$class
DefaultAssay(small) <- 'RNA' # let's try using the RNA slot first, but can also try the alra slot if we want
mark <- FindAllMarkers(small,logfc.threshold = 0.5,min.cells.feature = 1000) # Adding a min.cells.feature threshold here, to try to isolate signal that is common across datasets. Might not work as intended, but a reasonable first try
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
goi <- unique(mark$gene)
```

```{r}
small <- ScaleData(small)
small <- RunUMAP(small,features = goi)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

It's interesting that the RLMVEC group so nicely with the rest of the them. Otherwise though, there is major separation between starting cells and others. Let's see if we can get rid of this. One way to do so would be to remove the variance from these samples, like follows:

```{r}
Idents(small) <- small$orig.ident
sub <- subset(small,idents = c('FB13','FB14','BC1P3','BC1P6','MacAlv'),invert = T)
table(Idents(sub))
table(sub$class)
Idents(sub) <- sub$class
DefaultAssay(sub) <- 'RNA' # let's try using the RNA slot first, but can also try the alra slot if we want
mark <- FindAllMarkers(sub,logfc.threshold = 0.5,min.cells.feature = 1000) # Adding a min.cells.feature threshold here, to try to isolate signal that is common across datasets. Might not work as intended, but a reasonable first try
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
goi <- unique(mark$gene)
```

Now, we can use the above genes to embed all our data, but this time, hopefully it will have reduced the start vs. engineered displacement
```{r}
small <- RunUMAP(small,features = goi)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

Interesting. Mostly looks good, but the mesenchyme is still being cranky. Let's carve out the genes that are driving this separation of FB13 and FB14 from the main group(s):

```{r}
# We want to get rid of those features that distinguish FB13 and FB14...
Idents(small) <- small$orig.ident
annoying.genes <- FindMarkers(small,ident.1 = c('FB13','FB14'))
annoying.genes <- rownames(annoying.genes)
# ... but not those that distinguish the other fibroblasts from everything else
Idents(sub) <- sub$class # This object does not have FB13 or FB14 in it, so their variance will not be considered
good.genes <- FindMarkers(sub,ident.1 = 'Mesenchyme')
good.genes <- rownames(good.genes)
# which leaves only these bad actors
bad.genes <- annoying.genes[!(annoying.genes %in% good.genes)]
# removing these from the above goi to make goi.2
goi.2 <- goi[!(goi %in% bad.genes)]
# and trying again
small <- RunUMAP(small,features = goi.2)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

So stubborn! I'm beginning to think that there is some background RNA signal present in the engineered populations that is not present in the starting populations. This would make sense. What if we try to isolate and remove this signal?

```{r}
table(small$Dataset2)
```

Looks like we could try comparing Start to...everything else??

```{r}
Idents(small) <- small$Dataset2
annoying.2 <- FindMarkers(small,ident.1 = 'Start',min.pct = 0.2,min.cells.feature = 1000) # Want things that are not driven by one sample
annoying.2 <- rownames(annoying.2)
goi.3 <- goi.2[!(goi.2 %in% annoying.2)]
# and trying again
small <- RunUMAP(small,features = goi.3)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

I think this is good enough! Let's play with the UMAP parameters to make it look even prettier. This is a good guide for different things we could tune:
https://jlmelville.github.io/uwot/abparams.html

```{r}
small <- RunUMAP(small,features = goi.3,min.dist = 0.5)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

small <- RunUMAP(small,features = goi.3,min.dist = 1,repulsion.strength = 2)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

small <- RunUMAP(small,features = goi.3,min.dist = 1,repulsion.strength = 2,spread = 2)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# per https://jlmelville.github.io/uwot/abparams.html, let's try playing with a and b directly, instead of min.dist / spread

small <- RunUMAP(small,features = goi.3,a = 1,b=1)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

small <- RunUMAP(small,features = goi.3,a = 1,b=1.5)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

small <- RunUMAP(small,features = goi.3,a = 1.25,b=1.5)
p1 <- DimPlot(small,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(small,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

I think this last one, with features = goi.3,a = 1.25,b=1.5 looks pretty good for our purposes.

Let's see if it scales up nicely to the total 'downsampled' dataset that was used for NICHES edge creation. This will take a little longer, because we will be embedding many more cells, and we are not using principle components here, but rather 528 distinct genes/features, but I think it will still run in a reasonable time. Let's find out:

```{r}
downsampled <- RunUMAP(downsampled,features = goi.3,a = 1.25,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

Ok, this worked, and it is not terrible, but I'm not loving how separated each batch is, which I think will make the 'NICHES_FeaturePlots' a little difficult to interpret (I really want to be able to focus on the change in *representation* of edges, not *orientation* of edges. What if we bring down a and b, will it crush things together within each class a bit?

```{r}
downsampled <- RunUMAP(downsampled,features = goi.3,a = 1,b=1)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

Ok, if anything, that might have made things worse.

Next, let's try running the 'goi' calculations on the downsampled dataset. This may take a while...

```{r}
## 1. Find distinguishing genes, but not allowing it to look at the starting populations with issues
Idents(downsampled) <- downsampled$orig.ident
sub <- subset(downsampled,idents = c('FB13','FB14','BC1P3','BC1P6','MacAlv'),invert = T)
table(Idents(sub))
table(sub$class)
Idents(sub) <- sub$class
DefaultAssay(sub) <- 'RNA' # let's try using the RNA slot first, but can also try the alra slot if we want
mark <- FindAllMarkers(sub,logfc.threshold = 0.5,min.cells.feature = 1000) # Adding a min.cells.feature threshold here, to try to isolate signal that is common across datasets. Might not work as intended, but a reasonable first try
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
goi <- unique(mark$gene)

## 2. Specifically carve out genes which cause starting mesenchymal populations to segregate
# We want to get rid of those features that distinguish FB13 and FB14...
Idents(downsampled) <- downsampled$orig.ident
annoying.genes <- FindMarkers(downsampled,ident.1 = c('FB13','FB14'),logfc.threshold = 0.5,min.cells.feature = 1000) # Had to bump these thresholds up to make it not take 40 minutes
annoying.genes <- rownames(annoying.genes)
# ... but not those that distinguish the other fibroblasts from everything else
Idents(sub) <- sub$class # This object does not have FB13 or FB14 in it, so their variance will not be considered
good.genes <- FindMarkers(sub,ident.1 = 'Mesenchyme',logfc.threshold = 0.5,min.cells.feature = 1000)
good.genes <- rownames(good.genes)
# which leaves only these bad actors
bad.genes <- annoying.genes[!(annoying.genes %in% good.genes)]
# removing these from the above goi to make goi.2
goi.2 <- goi[!(goi %in% bad.genes)]

# 3. Specifically carve out genes which cause Start populations to segregate
Idents(downsampled) <- downsampled$Dataset2
annoying.2 <- FindMarkers(downsampled,ident.1 = 'Start',min.pct = 0.5,min.cells.feature = 5000) # Want things that are not driven by one sample
annoying.2 <- rownames(annoying.2)
goi.3 <- goi.2[!(goi.2 %in% annoying.2)]

## 4. Plotting
downsampled <- RunUMAP(downsampled,features = goi.3,a = 1.25,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

Hmm, that shift in the endothelium in Tri_E is bothersome. Remove?

```{r}
Idents(downsampled) <- downsampled$class
end <- subset(downsampled,idents = c('Endothelium'))
Idents(end) <- end$Dataset2
table(Idents(end))
icky <- FindMarkers(end,ident.1 = c('Tri_E'),min.pct = 0.5,logfc.threshold = 0.5,min.cells.feature = 400)
icky <- rownames(icky)
goi.4 <- goi.3[!(goi.3 %in% icky)]

downsampled <- RunUMAP(downsampled,features = goi.4,a = 1.25,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

```

Ok, that at least put the endothelium all together again. It's a start, not perfect yet.

What if we did the same thing that we just did for the endothelium also for the other classes that are showing annoying separation? Mes and Immune?

```{r}
# Remove Mesenchyme start vs. other variance
Idents(downsampled) <- downsampled$class
mes <- subset(downsampled,idents = c('Mesenchyme'))
Idents(mes) <- mes$Dataset2
table(Idents(mes))
icky.mes <- FindMarkers(mes,ident.1 = c('Start'),min.pct = 0.5,logfc.threshold = 0.5,min.cells.feature = 5000)
icky.mes <- rownames(icky.mes)
goi.5 <- goi.4[!(goi.4 %in% icky.mes)]

downsampled <- RunUMAP(downsampled,features = goi.5,a = 1.25,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

Ok, I'm calling this good enough for now on the feature-selection level. We can fine tune the embedding parameters directly if we like, to make the shape look prettier:

```{r}
# Toy with repulsion strength, whatever that means
downsampled <- RunUMAP(downsampled,features = goi.5,a = 1.25,b=1.5,repulsion.strength = 2)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# Bring up b, to try to make the subclusters blur together more (see https://jlmelville.github.io/uwot/abparams.html)
downsampled <- RunUMAP(downsampled,features = goi.5,a = 1.25,b=2,repulsion.strength = 2)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# Bring down a, to try to make the subclusters blur together more (see https://jlmelville.github.io/uwot/abparams.html)
downsampled <- RunUMAP(downsampled,features = goi.5,a = 0.8,b=1.5,repulsion.strength = 2)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# Return repulsion strength to the default (looks like it makes no difference, if we custom tune a and b. Good to know!)
downsampled <- RunUMAP(downsampled,features = goi.5,a = 0.8,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# bring down a even more?
downsampled <- RunUMAP(downsampled,features = goi.5,a = 0.5,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# try something from online?
downsampled <- RunUMAP(downsampled,features = goi.5,a = 0.1,b=1.9)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# try something from online?
downsampled <- RunUMAP(downsampled,features = goi.5,a = 0.01,b=0.9)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

# let's settle on this one, for now, cna be revised later
downsampled <- RunUMAP(downsampled,features = goi.5,a = 0.5,b=1.5)
p1 <- DimPlot(downsampled,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(downsampled,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

I think this is good enough for now, for further process development in this project. Let's save the object with its current embedding, for future reference, and the gene list to re-create if we feel like more parameter tuning is useful

```{r}
save(downsampled,file = 'downsampled.custom.embedding.2023-11-19.Robj')
save(goi.5,file = 'goi.5.2023-11-19.Robj')
```

(After going through all of this, I'm realizing, maybe we don't need the starter populations in this embedding at all! Maybe we only need to visualize the engineered organs for comparative connectomics?? The starting tissues will only have autocrine signaling anyways, so they will look like basically nothing in NICHES_FeaturePlots. Huh. To be revisited / discussed with Allie.)