---
title: "AllClass_Embedding_EngOnly_TriQuad_Dev"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10,fig.height = 7,warning = FALSE,message = FALSE)
```

Custom embedding for just Tri_L and Quad_E samples

#### Set up workspace
```{r}
# Require packages
require(ggplot2)
require(Seurat)
require(RColorBrewer)
require(cowplot)
require(dplyr)
library(scales)
library(viridis)

# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")

# Load color palette
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")
color_pals
```

#### Load the data from Part 1 that was used to calculate NICHES
```{r}
load('downsampled.imputed.Robj')
## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(downsampled$Final2)
```

## Reorder levels
```{r}
downsampled$class <- factor(downsampled$class,levels = c('Epithelium','Endothelium','Mesenchyme','Immune'))
```

As like last time, let's limit UMAP to only working with the variance distinguishing the different classes from each other, but this time, we will remove our starting populations at the outset.

### Remove starting populations
```{r}
Idents(downsampled) <- downsampled$Dataset2
table(Idents(downsampled))
sub <- subset(downsampled,idents=c('Tri_L','Quad_E'),invert=F)
table(Idents(sub))
table(sub$orig.ident)
```


Excellent. Now, we identify those genes that segregate classes, trying to get the calculation to not take very seriously those features that separate batches:

```{r}
table(sub$class)
Idents(sub) <- sub$class
DefaultAssay(sub) <- 'RNA'
# Old approach (not great here...)
mark <- FindAllMarkers(sub,min.pct = 0.75,logfc.threshold = 0.5,min.cells.feature = 1000,return.thresh = 0.0000000001) # These high threshold parameters are designed to (mostly) force it to only consider features common to multiple samples
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
goi <- unique(mark$gene)
goi

# new experiment for allie
Idents(sub) <- sub$orig.ident
bef12 <- subset(sub,idents = 'BEF12')
bef14 <- subset(sub,idents = 'BEF14')
bef15 <- subset(sub,idents = 'BEF15')
befm1 <- subset(sub,idents = 'BEFM1')
befm2 <- subset(sub,idents = 'BEFM2')
Idents(bef12) <- bef12$class
Idents(bef14) <- bef14$class
Idents(bef15) <- bef15$class
Idents(befm1) <- befm1$class
Idents(befm2) <- befm2$class
table(Idents(befm1))
table(Idents(befm2))
mark.bef12 <- FindAllMarkers(bef12,min.pct = 0.75,logfc.threshold = 1) # experimental
mark.bef14 <- FindAllMarkers(bef14,min.pct = 0.75,logfc.threshold = 1) # experimental
mark.bef15 <- FindAllMarkers(bef15,min.pct = 0.75,logfc.threshold = 1) # experimental
mark.befm1 <- FindAllMarkers(befm1,min.pct = 0.75,logfc.threshold = 1) # experimental
mark.befm2 <- FindAllMarkers(befm2,min.pct = 0.75,logfc.threshold = 1) # experimental

temp <- c(unique(mark.bef12$gene),unique(mark.bef14$gene),unique(mark.bef15$gene),unique(mark.befm1$gene),unique(mark.befm2$gene))
goi <- names(table(temp)>1)

goi

```

Seems reasonable as a starting point. Let's see what a default UMAP looks like.

```{r}
sub <- ScaleData(sub,features = goi)
sub <- RunUMAP(sub,features = goi,seed.use =104657) # playing with seed.use here will modify how the clusters are laid out on the page relative to one another
p1 <- DimPlot(sub,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(sub,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)

```

Not bad, but serious mesenchymal-specific separation (real signal). We don't care for this though, here. let's exclude the variance that separates the mesenchyme samples from one another:

```{r}
Idents(sub) <- sub$class
mes.sub <- subset(sub,idents = 'Mesenchyme')
Idents(mes.sub) <- mes.sub$orig.ident
mark.to.remove <- FindAllMarkers(mes.sub,min.pct = 0.5,logfc.threshold = 0.5)
goi.2 <- goi[!(goi %in% mark.to.remove$gene)]

# Add some features that I want in there (help endothelium to separate from mesenchyme)
goi.2 <- c(goi.2,'Cdh5')
mark.end <- FindMarkers(sub,ident.1 = 'Endothelium',min.pct = 0.5,logfc.threshold = 0.5)
mark.end$gene <- rownames(mark.end)
to.add <- mark.end$gene[!(mark.end$gene %in% mark.to.remove$gene)]
goi.3 <- unique(c(goi.2,to.add))

sub <- ScaleData(sub,features = goi.3)
sub <- RunUMAP(sub,features = goi.3,seed.use = 42,min.dist = 0.1) # playing with seed.use here will modify how the clusters are laid out on the page relative to one another
p1 <- DimPlot(sub,group.by = 'class',cols = color_pals$class_colors)
p2 <- DimPlot(sub,group.by = 'orig.ident',cols = color_pals$sample_colors,shuffle = T)
plot_grid(p1,p2,nrow=1)
```

Looks great!! LEt's save it


#### Save for later
```{r}
save(sub,file = 'downsampled.engOnly.TriQuad.customEmbed.2024-02-28.Robj') #hold on knit
```


#### Loading NICHES data

Previously computed in Part 1

```{r}
gc()
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
```

#### Subsetting to align with above phenotype object

Otherwise the above function might throw errors.

```{r}
ctc <- global.connectomics$CellToCell$alra 
Idents(ctc) <- ctc$Dataset2.Sending
table(Idents(ctc))
ctc.sub <- subset(ctc,idents = c('Tri_L','Quad_E'),invert=F)
table(Idents(ctc.sub))
ctc.sub <- ScaleData(ctc.sub)
table(ctc.sub$orig.ident)
```

## Fine tuning (removing the two macrophages in BEF12!)
```{r}
table(Idents(bef12))
barcodes.remove <- rownames(bef12@meta.data[which(bef12@meta.data$class == 'Immune'),])
sub.fine <- subset(sub,cells = barcodes.remove,invert=T)
table(Idents(sub.fine),sub.fine$orig.ident)

Idents(ctc.sub) <- ctc.sub$orig.ident
bef12.ctc <- subset(ctc.sub,idents = 'BEF12')
barcodes.remove.ctc <- rownames(bef12.ctc@meta.data[which(bef12.ctc@meta.data$class.Sending == 'Immune'),])
barcodes.remove.ctc <- c(barcodes.remove.ctc,
                         rownames(bef12.ctc@meta.data[which(bef12.ctc@meta.data$class.Receiving == 'Immune'),]))
ctc.sub.fine <- subset(ctc.sub,cells = barcodes.remove.ctc,invert=T)
table(Idents(ctc.sub.fine),ctc.sub.fine$class.Sending)
table(Idents(ctc.sub.fine),ctc.sub.fine$class.Receiving)

#save(sub.fine,file = 'transcriptome.TriQuad.finefilter.2024-02-29.Robj')
#save(ctc.sub.fine,file = 'connectome.TriQuad.finefilter.2024-02-29.Robj')
```

Now let's make network plots for specific selected mechanisms:

```{r}
# set up metadata
sub.fine$Condition <- sub.fine$Dataset2
ctc.sub.fine$Condition <- ctc.sub.fine$Dataset2.Sending
sub.fine$Condition <- factor(sub.fine$Condition,levels = c('Tri_L','Quad_E'))
ctc.sub.fine$Condition <- factor(ctc.sub.fine$Condition,levels = c('Tri_L','Quad_E'))

# fix color palette
color_pals$class <- color_pals$class_colors
# wd
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")

# set feature
MOI <- 'Wnt7a—Lrp6'

#plot
png(file = paste(MOI,'NetworkPlot.png',sep='_'),width = 6,height = 5,units = 'in',res=300)
NICHES_FeaturePlot(transcriptome.object = sub.fine,
                   connectome.object = ctc.sub.fine,
                   mechanism.of.interest = MOI,
                   legends.to.plot = 'class',
                   legend.palettes = color_pals,
                   split.by = 'Condition',
                   min.connectivity.thresh = 0.1,
                   connectivity.color.min = 0.1,
                   connectivity.color.max = 1,
                   line.thickness = 0.1,
                   black.points = T)
dev.off()

```

```{r}
# set feature
MOI <- 'Osm—Osmr'

NICHES_FeaturePlot(transcriptome.object = sub.fine,
                   connectome.object = ctc.sub.fine,
                   mechanism.of.interest = MOI,
                   legends.to.plot = 'class',
                   legend.palettes = color_pals,
                   split.by = 'Condition',
                   min.connectivity.thresh = 0.1,
                   connectivity.color.min = 0.1,
                   connectivity.color.max = 1,
                   line.thickness = 0.1)

```

```{r}
# set feature
MOI <- 'Wnt5a—Fzd7'

NICHES_FeaturePlot(transcriptome.object = sub.fine,
                   connectome.object = ctc.sub.fine,
                   mechanism.of.interest = MOI,
                   legends.to.plot = 'class',
                   legend.palettes = color_pals,
                   split.by = 'Condition',
                   min.connectivity.thresh = 0.1,
                   connectivity.color.min = 0.1,
                   connectivity.color.max = 1,
                   line.thickness = 0.1)

```

## for loop
```{r}
# define new metadata slot for finding maximum edge value
ctc.sub.fine$tempSlot <- paste(ctc.sub.fine$Dataset2.Sending,ctc.sub.fine$class.Joint)
temp <- AverageExpression(ctc.sub.fine,group.by = 'tempSlot',slot = 'data')
temp <- data.frame(temp$CellToCell)
temp$max <- apply(temp, 1, max, na.rm=TRUE)
temp$mean <- apply(temp, 1, mean, na.rm=TRUE)
temp$mid <- (temp$max+temp$mean)/2

# set threshold
min.connectivity.thresh <- 0.5

# epithelium
for(i in 1:length(row.labels.epi)){
feature = row.labels.epi[i]
png(file = paste(feature,'NetworkPlot.png',sep='_'),width = 6,height = 5,units = 'in',res=300)
print(NICHES_FeaturePlot(transcriptome.object = sub.fine,
                   connectome.object = ctc.sub.fine,
                   mechanism.of.interest = feature,
                   legends.to.plot = 'class',
                   legend.palettes = color_pals,
                   split.by = 'Condition',
                   min.connectivity.thresh = min.connectivity.thresh,
                   connectivity.color.min = log1p(min.connectivity.thresh),
                   connectivity.color.max = temp[feature,]$mid,
                   line.thickness = 0.1,
                   black.points = T))
dev.off()
}

# Endothelium (borrows from other script, need to clean up)
for(i in 1:length(row.labels.end)){
feature = row.labels.end[i]
png(file = paste(feature,'NetworkPlot.png',sep='_'),width = 6,height = 5,units = 'in',res=300)
print(NICHES_FeaturePlot(transcriptome.object = sub.fine,
                   connectome.object = ctc.sub.fine,
                   mechanism.of.interest = feature,
                   legends.to.plot = 'class',
                   legend.palettes = color_pals,
                   split.by = 'Condition',
                   min.connectivity.thresh = min.connectivity.thresh,
                   connectivity.color.min = log1p(min.connectivity.thresh),
                   connectivity.color.max = temp[feature,]$max,
                   line.thickness = 0.1,
                   black.points = T))
dev.off()
}
```



