---
title: "EndothelialReceivingWorking"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 20,fig.height = 10,warning = FALSE,message = FALSE)
```

#### Introduction:

Endothelial Receiving

#### Setup Packages


```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")

# Require packages
require(Seurat)
require(NICHES)
require(ggplot2)
require(cowplot)
require(dplyr)
require(knitr)

# Local functions
source('/Users/msbr/GitHub/NICHESMethods/R Scripts/ToolBuilding/NICHES_FeaturePlot.R')
```

#### Load NICHES data


```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
```


#### Subset out the just specific samples of interest, per Allie's request:

Here we limit all of our downstream analysis to information that came exclusively from the following 5 samples of interest, which will allow us to nicely isolate the effect of macrophages on mesenchymal to endothelial signaling.

```{r}
samples.of.interest <- c('BEF12','BEF14','BEF15','BEFM1','BEFM2')
# we will use this set of samples for downstream calculations in this workflow
```

#### Define colors

This is a temporary color palette, and we should consider updating / polishing for publication.

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")

## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(global.connectomics$CellToCell$alra$SendingType)
```

#### Isolate Endothelial Receiving edges

```{r}
temp <- global.connectomics$CellToCell$alra # we will use the imputed data, for now
Idents(temp) <- temp$class.Receiving
table(Idents(temp))
end.rec <- subset(temp, idents = 'Endothelium')
Idents(end.rec) <- end.rec$orig.ident
table(Idents(end.rec))
end.rec <- subset(end.rec,idents = c(samples.of.interest))
table(Idents(end.rec))
```


#### Endo Receiving

```{r}
end.rec <- ScaleData(end.rec)
end.rec <- FindVariableFeatures(end.rec)
end.rec <- RunPCA(end.rec,npcs = 100)
ElbowPlot(end.rec,ndims = 100)
end.rec <- RunUMAP(end.rec,dims = 1:25)
DimPlot(end.rec)

p1 <- DimPlot(end.rec,group.by = 'Dataset2.Sending',cols=color_pals$dataset_colors,raster=F,shuffle = T)
p2 <- DimPlot(end.rec,group.by = 'orig.ident',cols = color_pals$sample_colors,raster=F,shuffle = T,label=T)
p3 <- DimPlot(end.rec,group.by = 'class.Sending',cols = color_pals$class_colors,raster=F,shuffle = T)
p4 <- DimPlot(end.rec,group.by = 'SendingType',cols = color_pals$type_colors,raster=F,shuffle = T)
p5 <- DimPlot(end.rec,group.by = 'ReceivingType',cols = color_pals$type_colors,raster=F,shuffle = T)
plot_grid(p1,p2,p3,p4,p5)

png(file='EndotheliumReceiving.UMAPS.png',width=20,height=12,units = 'in',res=300)
print(plot_grid(p1,p2,p3,p4,p5))
dev.off()
```

#### Cluster into archetypes

```{r}
end.rec <- FindNeighbors(end.rec,dims = 1:25)
end.rec <- FindClusters(end.rec,resolution = 0.2)
DimPlot(end.rec,label=T)

p6 <- DimPlot(end.rec,label=T)

png(file='EndotheliumReceivingClustered.UMAPS.png',width=20,height=12,units = 'in',res=300)
print(plot_grid(p1,p2,p3,p4,p5,p6))
dev.off()
```

QUAD CULTURE PROMOTES ANGIOGENESIS

Cluster 6 is that brought by macrophages
4,5 and 8 are associated with cycling cells AND Quad Culture

```{r}
mark <- FindAllMarkers(end.rec,min.pct = 0.1,logfc.threshold = 0.1,only.pos=T)
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
kable(mark %>% group_by(cluster) %>% top_n(20,ratio))
```
```{r}
end.rec <- RenameIdents(end.rec,
                        '0'='Rspondins / Vegfa / Homeostasis',
                        '1'='BEFM1',
                        '2'='Endothelial Autocrine',
                        '3'='Barrier?',
                        '4'='Mesenchymal Angiogenic',
                        '5'='Epithelial Angiogenic',
                        '6'='Immune',
                        '7'='BEFM1',
                        '8'='Endothelial Angiogenic')

```
How how how is the top markers Rspo3-Lgr5
>FeaturePlot(end.rec,'Wnt7a—Fzd9',label=T,max.cutoff = 1,order=T)




#### Load transcriptomic data (not NICHES data) so that we can make NICHESFeaturePlots (which leverage BOTH data types at once)

```{r}
 # Transcriptome object is called 'sub'
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/downsampled.engOnly.customEmbed.2023-12-14.Robj") # 'sub'
```

#### Test that the plotting function(s) work

```{r}

# First we have to scale the connectivity data, just in case we haven't before
connectivity <- global.connectomics[['CellToCell']][['alra']]
connectivity <- ScaleData(connectivity)

# Then we set our mechanism of interest for plotting (from the mark test above)
MOI <- 'Wnt7a—Lrp6'

# Set up colors for legend plots
legend.palettes <- color_pals
# Make the names of legend.palettes correlates with the legends requested below
names(legend.palettes)[3] <- 'class'
names(legend.palettes)[1] <- 'Sample' 

# Make a network plot with one legend, with plots side by side:
network.plot <- NICHES_FeaturePlot(transcriptome.object = sub,
                                   connectome.object = connectivity,
                                   mechanism.of.interest = MOI,
                                   legends.to.plot = 'class',
                                   transpose = TRUE,
                                   legend.palettes = legend.palettes)
network.plot

# Make a network plot with two legends, with plots on top of one another:
network.plot <- NICHES_FeaturePlot(transcriptome.object = sub,
                                   connectome.object = connectivity,
                                   mechanism.of.interest = MOI,
                                   legends.to.plot = c('class','Sample'),
                                   transpose = FALSE,
                                   legend.palettes = legend.palettes)
network.plot

```

Good. The plotting function works. But we want to split by Mono / Co / Tri / Quad. Currently we don't have a metadata slot that groups in this way. So, let's first create the appropriate metadata slot and organize the levels appropriately for plotting downstream. Note that we need to create parallel metadat slots in both the transcriptomic and connectomic objects.

```{r}
# Build a new metadata slot ('Dataset4') to split based on M/C/T/Q
Idents(sub) <- sub$Dataset2
Idents(connectivity) <- connectivity$Dataset2.Sending
table(Idents(sub))
table(Idents(connectivity))
sub <- RenameIdents(sub,
                       'Tri_E'='Tri',
                       'Tri_L'='Tri',
                       'Quad_E'='Quad',
                       'Quad_L'='Quad')
connectivity <- RenameIdents(connectivity,
                             'Tri_E'='Tri',
                             'Tri_L'='Tri',
                             'Quad_E'='Quad',
                             'Quad_L'='Quad')
sub$Dataset4 <- Idents(sub)
connectivity$Dataset4 <- Idents(connectivity)

sub$Dataset4 <- factor(sub$Dataset4,levels=c('Mono','Co','Tri','Quad'))
connectivity$Dataset4 <- factor(connectivity$Dataset4,levels=c('Mono','Co','Tri','Quad'))
```

Great. Now we are prepared to make a NICHES_FeaturePlot that splits by 'Dataset4'

```{r}
# Split according to this new slot
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class','Sample'),
                       transpose = FALSE,
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```

And we can experiment further:
```{r}
MOI <- 'Osm—Osmr' # change MOI, picked from markers for cluster 8 above
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class'), # split only by class
                       transpose = TRUE, # transpose
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```

This is very cool. Sounds like there is good evidence for macrophage-derived OSM reprogramming fibroblasts so that they help our other cells grow and migrate, in pancreatic cancer at least: https://pubmed.ncbi.nlm.nih.gov/34921158/

```{r}
MOI <- 'Tgfb1—Itgb6' # change MOI, picked from markers for cluster 7 above
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class'), # split only by class
                       transpose = TRUE, # transpose
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```
```{r}
MOI <- 'Tnf—Ltbr' # change MOI, picked from markers for cluster 6 above
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class'), # split only by class
                       transpose = TRUE, # transpose
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```
DELIGHTFUL! So, it turns out that adding macrophages to our system makes the ENDOTHELIUM and Endothelium and a little bit of the EPITHELIUM secrete TNF which can be sensed by Macrophages and Epithelial cells.

What about matrix production?

```{r}
MOI <- 'Npnt—Itgb1' # change MOI, picked from markers for cluster 6 above (epithelial origin)

NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class'), # split only by class
                       transpose = TRUE, # transpose
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```

Very cool. NPNT is a matrix protein that promotes angiogenesis: https://www.nature.com/articles/srep36210

Also:
```{r}
MOI <- 'Bmp3—Bmpr1b' # change MOI, picked from markers for cluster 6 above (epithelial origin)

NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class'), # split only by class
                       transpose = TRUE, # transpose
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```

Not sure what Bmp3 does in the lung...

And:
```{r}
MOI <- 'Angpt1—Tie1' # change MOI, picked from markers for cluster 4 above (mesenchymal origin)

NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class'), # split only by class
                       transpose = TRUE, # transpose
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.5,
                       legend.alpha = 0.75) # bump the legend alpha up
```

Angpt1—Tie1 is a major pro-homeostatic microvascular stabilizing interaction present in the alveolus during normal homeostasis from fibroblasts to endothelium. It is strongly upregulated in quad-culture.

#### Conclusions:

I think this is an efficient and practical workflow to analyze our entire dataset. It means we only have to do this three more times to feel confident that we have at least done a reasonable observation of 'everything' and aren't missing anything obvious. It runs quick, is generalized so it should be easily adaptible to the other classes, and we can generate exciting results and feel confidenat about the answers! I'm on board and I think this works well. 



