---
title: "EndothelialReceivingWorking"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 24,fig.height = 22,warning = FALSE,message = FALSE)
```

## Introduction

This script analyzes endothelial niche signaling in the BEFM project

## Set WD
```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")
```

## Packages
```{r}
require(Seurat)
require(NICHES)
require(dplyr)
require(ggthemes)
require(ggplot2)
require(cowplot)
require(knitr)
```


# Local functions
```{r}
source('/Users/msbr/GitHub/NICHESMethods/R Scripts/ToolBuilding/NICHES_FeaturePlot.R')
```

#### Load saved NICHES data from Part 1

Note that no data filtration has been performed, per decisions in Part 2.

As a reminder, these data were made from a 5000 cell-per-sample downsampled object that is then fed into NICHES and the outputs saved. 

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
```

## Inspect
```{r}
global.connectomics
```
#### Subset out the just specific samples of interest, per Allie's request:

Here we limit all of our downstream analysis to information that came exclusively from the following 5 samples of interest, which will allow us to nicely isolate the effect of macrophages on mesenchymal to endothelial signaling.

```{r}
samples.of.interest <- c('BEF12','BEF14','BEF15','BEFM1','BEFM2')
# we will use this set of samples for downstream calculations in this workflow
```


#### Define colors

This is a temporary color palette, and we should consider updating / polishing for publication.

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")

## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(global.connectomics$CellToCell$alra$SendingType)
```

#### Isolate Endothelial Receiving edges

```{r}
temp <- global.connectomics$CellToCell$alra # we will use the imputed data, for now
Idents(temp) <- temp$class.Receiving
table(Idents(temp))
end.rec <- subset(temp, idents = 'Endothelium')
Idents(end.rec) <- end.rec$orig.ident
table(Idents(end.rec))
end.rec <- subset(end.rec,idents = c(samples.of.interest))
table(Idents(end.rec))
```
## Clean

Since we are no longer prioritizing BCL5, we can clean these data to help us make prettier plots

```{r}
VlnPlot(end.rec,'nFeature_CellToCell',group.by = 'orig.ident')
end.rec <- subset(end.rec,subset = nFeature_CellToCell>100)
VlnPlot(end.rec,'nFeature_CellToCell',group.by = 'orig.ident')
```

Excellent. Let's take a global look at this edge category before moving further:

#### Endothelium Receiving - Embed and Cluster

```{r}
end.rec <- ScaleData(end.rec)
end.rec <- FindVariableFeatures(end.rec)
end.rec <- RunPCA(end.rec,npcs = 50)
ElbowPlot(end.rec,ndims = 50)
end.rec <- RunUMAP(end.rec,dims = 1:10)
end.rec <- FindNeighbors(end.rec,dims = 1:10)
end.rec <- FindClusters(end.rec,resolution = 0.1)
```

## Endo Plot
```{r}
p1 <- DimPlot(end.rec,group.by = 'class.Sending',shuffle = T)+ggtitle('Sending CellClass')
p2 <- DimPlot(end.rec,group.by = 'class.Receiving',shuffle = T)+ggtitle('Receiving CellClass')
p3 <- DimPlot(end.rec,group.by = 'Dataset2.Sending',shuffle = T)+ggtitle('Dataset2')
p4 <- DimPlot(end.rec,group.by = 'SendingType',shuffle = T)+ggtitle('Sending CellType')
p5 <- DimPlot(end.rec,group.by = 'ReceivingType',shuffle = T)+ggtitle('Receiving CellType')
p6 <- DimPlot(end.rec,group.by = 'orig.ident',shuffle = T)+ggtitle('Sample')

# Define plot blocks
small.mult <- cowplot::plot_grid(p1,p2,p3,p4,p5,p6,ncol = 3)
sig.arch <- DimPlot(end.rec,group.by = 'seurat_clusters',shuffle = T,label = T)+ggtitle('Signaling Archetypes')
split.arch <- DimPlot(end.rec,split.by = 'orig.ident',group.by = 'seurat_clusters',shuffle = T,ncol = 2)+ggtitle('Signaling Archetypes (Split By Sample)')

# Assemble blocks in cowplot
top <- small.mult
bottom <- cowplot::plot_grid(sig.arch,split.arch,rel_widths = c(1.2,1))
plot <- cowplot::plot_grid(top,bottom,nrow = 2)
plot

png('Endothelial_Receiving.png',width=24,height = 22,units = 'in',res=300)
print(plot)
dev.off()
```


## Save for later
```{r}
save(end.rec,file = 'end.rec.Robj')
```
