---
title: "EpiRecFeatures"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 24,fig.height = 22,warning = FALSE,message = FALSE)
```

## Introduction

This script identifies specific features of interest, landing on epithelium, in the BEFM project.

## Set WD
```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")
```

## Packages
```{r}
require(Seurat)
require(NICHES)
require(dplyr)
require(ggthemes)
require(ggplot2)
require(cowplot)
require(knitr)
```

## Color palette

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")
color_pals
```

## Load Epi Rec from save
```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/epi.rec.Robj")
```

## Inspect
```{r}
epi.rec
```

## Update to v5
```{R}
epi.rec <- UpdateSeuratObject(epi.rec)
epi.rec
```
## Dimplot for review
```{r}
p0 <- DimPlot(
  epi.rec,
  reduction = "umap",
  group.by = c('class.Sending',"Dataset2.Sending", "orig.ident", "seurat_clusters"),
  combine = FALSE, label.size = 8,shuffle = T,label=T)
p0 <- cowplot::plot_grid(plotlist = p0,nrow = 1)
p0
```
## Define archetypes for analysis
```{r}
epi.rec$Archetype <- epi.rec$seurat_clusters
Idents(epi.rec) <- epi.rec$Archetype
```

## Dimplot for review
```{r}
p0 <- DimPlot(
  epi.rec,
  reduction = "umap",
  group.by = c('Archetype','class.Sending',"Dataset2.Sending", "orig.ident", "pca_clusters"),
  combine = FALSE, label.size = 8,shuffle = T,label=T)
p0 <- cowplot::plot_grid(plotlist = p0,nrow = 1)
p0
```

Cluster 3 and 4 will be particularly of interest

3 = Epithelial autocrine associated with BEFM1
4 = Mesenchymal - Epithelial associated with BEFM1

# Differential expression across archetype

```{r}
mark <- FindAllMarkers(epi.rec,min.pct = 0.1,logfc.threshold = 0.1)
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
kable(mark %>% group_by(cluster) %>% top_n(20,ratio))
```

## Cluster 0 (Mes-Epi Common)
```{r}
goi.0 <-  c('Dlk1—Notch1','Bmp5—Bmpr1b','Rspo3—Lrp6',
                         'Angpt1—Itgb1','Bmp5—Acvr2a','Thbs2—Itgb1',
                         'Bmp5—Bmpr1a','Thbs2—Cd47','Ptn—Sdc1','Mst1—Mst1r','Fgf7—Fgfr1','Igf2—Igf2r','Bmp5—Bmpr1b','Ptn—Sdc1','Fgf10—Fgfr2','Dkk2—Lrp6','Bmp5—Bmpr2')
FeaturePlot(epi.rec,
            features = goi.0,order = T,pt.size=1,label=T,label.size = 8)

```

## Cluster 1 (Epi-Autocrine Common)
```{r}
goi.1 <-  c('Efna3—Epha1','Efna3—Epha2','Efna3—Epha4','Ntf4—Ngfr','Il10—Il10rb','Fgf9—Fgfr1','Dlk2—Notch1','Wnt3a—Fzd6','Wnt3a—Lrp6','Wnt3a—Ryk','Fgf9—Fgfr2','Bmp7—Eng','Bmp7—Bmpr1b','Wnt7b—Fzd1','Bmp7—Bmpr2','Areg—Erbb3','Ereg—Erbb3','Areg—Egfr')
FeaturePlot(epi.rec,
            features = goi.1,order = T,pt.size=1,label=T,label.size = 8)

```

## Cluster 2 (End-Epi Common)
```{r}
goi.2 <-  c('Dll1—Notch2','Dll1—Notch1','Jag1—Notch4','Dll1—Notch3','Bmp3—Bmpr1b','Dll1—Notch4','Gas6—Axl','Tnfsf12—Tnfrsf25','Jag1—Notch3','Bmp6—Bmpr2','Bmp3—Bmpr2','Vegfc—Nrp2','Vip—Sctr','Tnfsf4—Tnfrsf4','Dll4—Notch1','Dll4—Notch2','Pdgfb—Pdgfrb','Angpt2—Tie1','Dll4—Notch3','Dll4—Notch4','Dhh—Ptch1','Sema6d—Plxna1')
FeaturePlot(epi.rec,
            features = goi.2,order = T,pt.size=1,label=T,label.size = 8)

```

## Cluster 3 (Epi-Autocrine Quad)
```{r}
goi.3 <- c('Tnf—Tnfrsf1b','Sct—Sctr','Wnt7a—Lrp6',
                         'Npnt—Itgb1','Ereg—Egfr','Wnt7b—Lrp5',
                         'Hbegf—Egfr','Lamb3—Itga3','Npnt—Itgb1','Tnf—Tnfrsf21','Tnf—Tnfrsf21','Tnf—Ltbr','Lamb3—Itga2','Mmp7—Cd44','Lamb3—Itga2','Btc—Egfr')
FeaturePlot(epi.rec,
            features = goi.3,
            order = T,pt.size = 1,label=T,label.size = 8)
```

## Cluster 4 (Mes-Epi Quad)
```{r}
goi.4 <- c('Fgf1—Fgfr3','Fgf1—Egfr','Fgf1—Fgfr2','Fgf2—Fgfr3','Fgf12—Fgfr3','Lama1—Itga2',
                         'Csf2—Csf2ra','Ngf—Maged1','Wnt16—Fzd6',
                         'Fgf7—Fgfr3','Fn1—Itgb6','Vegfa—Ephb2',
                         'Wnt5a—Lrp5','Col4a1—Itga2','Tnc—Egfr','Tnc—Egfr')
  
FeaturePlot(epi.rec,
            features = goi.4,
            order = T,pt.size = 1,label=T,label.size = 8)
```

## Cluster 5 (Imm-Epi Quad)
```{r}
goi.5 <- c('Osm—Osmr','Il1b—Il1r1','Il1b—Adrb2','Tnfsf15—Tnfrsf25','Ebi3—Il6st','Il18—Il18r1','Camp—Igf1r','Tnfsf13—Tnfrsf1a','Nrg4—Egfr','Nrg4—Erbb2','Efna2—Epha1','Wnt5a—Ryk','Amh—Egfr','Hgf—Met','Igf1—Igf1r','Cxcl2—Cxcr2')
  
FeaturePlot(epi.rec,
            features = goi.5,
            order = T,pt.size = 1,label=T,label.size = 8)
```

## Heatmap
```{r}
goi <- mark %>% group_by(cluster) %>% top_n(100,power)
#goi <- c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5)
png('EpiRec.Heatmap.png',width=15,height=24,units='in',res=300)
CustomHeatmap(epi.rec,
              data.type = 'CellToCell',
                 primary = 'seurat_clusters' ,
                 secondary = 'Dataset2.Sending' ,
                 tertiary = 'class.Sending' ,
                 quarternary = 'orig.ident' ,
                 primary.cols = NULL,
                 secondary.cols = color_pals$dataset_colors, # Need to be a named list of colors
                 tertiary.cols = color_pals$class_colors,
                 quarternary.cols = color_pals$sample_colors,
                 features = goi$gene,
                  #features = goi,
                 labels = c('Signaling Archetype','Condition','Sending Cell Class','Sample'),
                  selected.row.anotations=c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5),
              selected.label.size = 10)
dev.off()
```

```{r}
goi <- mark %>% group_by(cluster) %>% top_n(100,power)
#goi <- c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5)
png('EpiRec.Heatmap.2.png',width=15,height=24,units='in',res=300)
CustomHeatmap(epi.rec,
              data.type = 'CellToCell',
                 primary = 'class.Sending' ,
                 secondary = 'Dataset2.Sending' ,
                 tertiary = 'seurat_clusters' ,
                 quarternary = 'orig.ident' ,
                 primary.cols = color_pals$class_colors,
                 secondary.cols = color_pals$dataset_colors, # Need to be a named list of colors
                 tertiary.cols = NULL,
                 quarternary.cols = color_pals$sample_colors,
                 features = goi$gene,
                  #features = goi,
                 labels = c('Sending Cell Class','Condition','Signaling Archetype','Sample'),
                  selected.row.anotations=c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5),
              selected.label.size = 10)
dev.off()
```

```{r}
goi <- mark %>% group_by(cluster) %>% top_n(100,power)
#goi <- c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5)
secondary.colors <- c('#6E2594','#F7C59F')
names(secondary.colors) <- c('Tri_E','Quad_E')
quarternary.colors <-  c('#55DDE0','#33658A','#2F4858','#F6AE2D','#F26419')
names(quarternary.colors) <- c('BEF12','BEF14','BEF15','BEFM2','BEFM1')
png('EpiRec.Heatmap.3.png',width=15,height=24,units='in',res=300)
CustomHeatmap(epi.rec,
              data.type = 'CellToCell',
                 primary = 'seurat_clusters' ,
                 secondary = 'Dataset2.Sending' ,
                 tertiary = 'class.Sending' ,
                 quarternary = 'orig.ident' ,
                 primary.cols = NULL ,
                 secondary.cols = secondary.colors, # Need to be a named list of co lors
                 tertiary.cols = color_pals$class_colors,
                 quarternary.cols = quarternary.colors,
                 features = goi$gene,
                  #features = goi,
                 labels = c('Sending Cell Class','Condition','Signaling Archetype','Sample'),
                  selected.row.anotations=c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5),
              selected.label.size = 10)
dev.off()
```

```{r}
#goi <- mark %>% group_by(cluster) %>% top_n(100,power)
goi <- c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5)
row.labels <- goi[grep('Wnt|Rspo|Lgr|Lrp|Fzd|Ryk',goi)]

secondary.colors <- c('#6E2594','#F7C59F')
names(secondary.colors) <- c('Quad_E','Tri_L')
quarternary.colors <-  c('#55DDE0','#33658A','#2F4858','#F6AE2D','#F26419')
names(quarternary.colors) <- c('BEF12','BEF14','BEF15','BEFM2','BEFM1')


png('EpiRec.Heatmap.4.png',width=15,height=15,units='in',res=300)
CustomHeatmap(epi.rec,
              data.type = 'CellToCell',
                 primary = 'seurat_clusters' ,
                 secondary = 'Dataset2.Sending' ,
                 tertiary = 'class.Sending' ,
                 quarternary = 'orig.ident' ,
                 primary.cols = NULL ,
                 secondary.cols = secondary.colors, # Need to be a named list of co lors
                 tertiary.cols = color_pals$class_colors,
                 quarternary.cols = quarternary.colors,
                 #features = goi$gene,
                  features = goi,
                 labels = c('Sending Cell Class','Condition','Signaling Archetype','Sample'),
                  selected.row.anotations=row.labels,
              selected.label.size = 10)
dev.off()
```

```{r}
#goi <- mark %>% group_by(cluster) %>% top_n(100,power)
goi <- c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5)
row.labels <- goi[grep('Wnt|Rspo|Lgr|Lrp|Fzd|Ryk',goi)]


epi.rec$class.Sending <- factor(epi.rec$class.Sending,
                                   levels = c('Epithelium', 'Endothelium', 'Mesenchyme', 'Immune'))
epi.rec$Dataset2.Sending <- factor(epi.rec$Dataset2.Sending,
                                   levels = c('Tri_L','Quad_E'))

png('EpiRec.Heatmap.5.png',width=10,height=10,units='in',res=300)
CustomHeatmap(epi.rec,
              data.type = 'CellToCell',
                 primary = 'Archetype' ,
                 secondary = 'class.Sending' ,
                 tertiary = 'Dataset2.Sending' ,
                 quarternary = 'orig.ident' ,
                 primary.cols = NULL ,
                 secondary.cols = color_pals$class_colors,
                 tertiary.cols = color_pals$dataset_colors[5:6],
                 quarternary.cols = color_pals$sample_colors,
                 #features = goi$gene,
                  features = goi,
                 labels = c('Signaling Archetype','Sending Cell Class','Condition','Sample'),
                  selected.row.anotations=row.labels,
              selected.label.size = 10)
dev.off()
```
```{r}
Idents(epi.rec) <- epi.rec$seurat_clusters
table(Idents(epi.rec))
epi.rec <- RenameIdents(epi.rec,
                        '0'='Baseline Mes-Epi',
                        '1'='Baseline Epi-Auto',
                        '2'='Baseline Endo-Epi',
                        '3'='Regenerative Epi-Auto',
                        '4'='Regenerative Mes-Epi',
                        '5'='Macrophage-Epi')
table(Idents(epi.rec))
epi.rec$Archetype <- Idents(epi.rec)
```
## Reorder levels
```{r}
epi.rec$class.Sending <- factor(epi.rec$class.Sending,levels = c('Epithelium','Endothelium','Mesenchyme','Immune'))
```

## scale
```{r}
epi.rec <- ScaleData(epi.rec)
```
## Final heatmap for jalenlia poster
```{r}
goi.0 <-  c('Angpt1—Itgb1','Bmp5—Acvr2a','Thbs2—Itgb1',
                         'Fgf10—Fgfr2','Bmp5—Bmpr1a','Thbs2—Cd47','Dlk1—Notch1','Bmp5—Bmpr1b','Rspo3—Lrp6',
                         'Angpt1—Itgb1','Bmp5—Acvr2a','Thbs2—Itgb1',
                         'Bmp5—Bmpr1a','Thbs2—Cd47','Ptn—Sdc1','Mst1—Mst1r','Fgf7—Fgfr1','Igf2—Igf2r','Bmp5—Bmpr1b','Ptn—Sdc1','Dkk2—Lrp6','Bmp5—Bmpr2')

goi.1 <-  c('Efna3—Epha1','Efna3—Epha2','Efna3—Epha4','Ntf4—Ngfr','Il10—Il10rb','Fgf9—Fgfr1','Dlk2—Notch1','Wnt3a—Fzd6','Wnt3a—Lrp6','Wnt3a—Ryk','Wnt7b—Fzd1','Fgf9—Fgfr2','Bmp7—Eng','Bmp7—Bmpr1b','Bmp7—Bmpr2','Areg—Erbb3','Ereg—Erbb3','Areg—Egfr')

goi.2 <-  c('Jag1—Notch4','Dll1—Notch3','Bmp3—Bmpr1b','Dll1—Notch4','Gas6—Axl','Tnfsf12—Tnfrsf25','Jag1—Notch3','Bmp6—Bmpr2','Bmp3—Bmpr2','Vegfc—Nrp2','Vip—Sctr','Dll1—Notch2','Dll1—Notch1','Dhh—Ptch1','Tnfsf4—Tnfrsf4','Dll4—Notch1','Dll4—Notch2','Pdgfb—Pdgfrb','Angpt2—Tie1','Dll4—Notch3','Dll4—Notch4','Sema6d—Plxna1')

goi.3 <- c('Tnf—Tnfrsf1b','Sct—Sctr','Wnt7a—Lrp6',
                         'Npnt—Itgb1','Ereg—Egfr','Wnt7b—Lrp5',
                         'Hbegf—Egfr','Lamb3—Itga3','Npnt—Itgb1','Tnf—Tnfrsf21','Tnf—Tnfrsf21','Tnf—Ltbr','Lamb3—Itga2','Mmp7—Cd44','Lamb3—Itga2','Btc—Egfr')

goi.4 <- c('Fgf1—Fgfr3','Fgf1—Egfr','Fgf1—Fgfr2','Fgf2—Fgfr3','Fgf12—Fgfr3','Lama1—Itga2',
                         'Csf2—Csf2ra','Ngf—Maged1','Wnt16—Fzd6',
                         'Fgf7—Fgfr3','Fn1—Itgb6','Vegfa—Ephb2',
                         'Wnt5a—Lrp5','Wnt5a—Ryk','Col4a1—Itga2','Tnc—Egfr')
  
goi.5 <- c('Amh—Egfr','Hgf—Met','Igf1—Igf1r','Cxcl2—Cxcr2','Ebi3—Il6st','Il18—Il18r1','Tnfsf13—Tnfrsf1a','Nrg4—Egfr','Nrg4—Erbb2','Efna2—Epha1')
  
goi.epi <- unique(c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5))

row.labels.epi <- c('Bmp5—Bmpr1b','Rspo3—Lrp6','Fgf10—Fgfr2',
                'Il10—Il10rb','Wnt3a—Fzd6','Wnt7b—Fzd1',
                'Dll1—Notch2','Vip—Sctr','Dhh—Ptch1',
                'Sct—Sctr','Wnt7a—Lrp6','Wnt7b—Lrp5',"Lamb3—Itga3",'Tnf—Ltbr',
                'Fgf7—Fgfr3','Fn1—Itgb6','Col4a1—Itga2','Wnt5a—Ryk' ,'Il18—Il18r1','Nrg4—Egfr')

png('EpiRec.Heatmap.png',width=7,height=9,units='in',res=300)
CustomHeatmapOnly3(epi.rec,
              data.type = 'CellToCell',
                 primary = 'Archetype' ,
                 secondary = 'class.Sending' ,
                 tertiary = 'orig.ident' ,
                 #quarternary = 'orig.ident' ,
                 primary.cols = NULL ,
                 secondary.cols = color_pals$class_colors,
                 tertiary.cols = color_pals$sample_colors,#color_pals$dataset_colors[5:6],
                 #quarternary.cols = color_pals$sample_colors,
                 features = goi.epi,
                 labels = c('Signaling Archetype','Sending Cell Class','Sample'),
                selected.row.anotations=row.labels.epi,
              selected.label.size = 10,
              use.scale.data = T,
              range.frac = 0.2)
dev.off()
```

### Load and pre-process data for Circuit Plot, temporary for Janelia poster
```{r}
# test on allie's dataset
# connectivity data
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
# transcriptomic data
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/downsampled.engOnly.customEmbed.2023-12-14.Robj") # 'sub'
# First we have to scale the connectivity data, just in case we haven't before
connectivity <- global.connectomics[['CellToCell']][['alra']]
connectivity <- ScaleData(connectivity)
# Build a new metadata slot ('Dataset4') to split based on M/C/T/Q
Idents(sub) <- sub$Dataset2
Idents(connectivity) <- connectivity$Dataset2.Sending
table(Idents(sub))
table(Idents(connectivity))
sub <- RenameIdents(sub,
                    'Tri_E'='Tri',
                    'Tri_L'='Tri',
                    'Quad_E'='Quad',
                    'Quad_L'='Quad')
connectivity <- RenameIdents(connectivity,
                             'Tri_E'='Tri',
                             'Tri_L'='Tri',
                             'Quad_E'='Quad',
                             'Quad_L'='Quad')
sub$Dataset4 <- Idents(sub)
connectivity$Dataset4 <- Idents(connectivity)

sub$Dataset4 <- factor(sub$Dataset4,levels=c('Mono','Co','Tri','Quad'))
connectivity$Dataset4 <- factor(connectivity$Dataset4,levels=c('Mono','Co','Tri','Quad'))

# Divide up into 4 systems
Idents(sub) <- sub$Dataset4
tran.mono <- subset(sub,idents = 'Mono')
tran.co <- subset(sub,idents = 'Co')
tran.tri <- subset(sub,idents = 'Tri')
tran.quad <- subset(sub,idents = 'Quad')
Idents(connectivity) <- connectivity$Dataset4
con.mono <- subset(connectivity,idents = 'Mono')
con.co <- subset(connectivity,idents = 'Co')
con.tri <- subset(connectivity,idents = 'Tri')
con.quad <- subset(connectivity,idents = 'Quad')

# Adjust the tri objects to get rid of the tiny population of contaminating macrophages
table(tran.tri$class)
table(con.tri$class.Sending)
table(con.tri$class.Receiving)

Idents(tran.tri) <- tran.tri$class
tran.tri <- subset(tran.tri,idents = 'Immune',invert = T)

Idents(con.tri) <- con.tri$class.Sending
con.tri <- subset(con.tri,idents = 'Immune',invert = T)
Idents(con.tri) <- con.tri$class.Receiving
con.tri <- subset(con.tri,idents = 'Immune',invert = T)

table(tran.tri$class)
table(con.tri$class.Sending)
table(con.tri$class.Receiving)

# defin temp plotting function
TempFunc <- function(feature,
                     max.edge.value = 0.1){
  p4 <- CircuitPlot(transcr.obj = tran.quad,
              connect.obj = con.quad,
              feature = feature,
              group.by = 'class',
              edge.fixed.size = T,min.edge.value = 0,max.edge.value = max.edge.value,cols.use = cols.use)
  p3 <- CircuitPlot(transcr.obj = tran.tri,
              connect.obj = con.tri,
              feature = feature,
              group.by = 'class',
              edge.fixed.size = T,min.edge.value = 0,max.edge.value = max.edge.value,cols.use = cols.use)
  # p2 <- CircuitPlot(transcr.obj = tran.co,
  #             connect.obj = con.co,
  #             feature = feature,
  #             group.by = 'class',
  #             edge.fixed.size = T,min.edge.value = 0,max.edge.value = max.edge.value,cols.use = cols.use)
  # p1 <- CircuitPlot(transcr.obj = tran.mono,
  #             connect.obj = con.mono,
  #             feature = feature,
  #             group.by = 'class',
  #             edge.fixed.size = T,min.edge.value = 0,max.edge.value = max.edge.value,cols.use = cols.use)
  print(cowplot::plot_grid(p3,p4,nrow = 1))
}

```

## Circuitplots for janelia
```{r}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")
cols.use <- color_pals$class_colors

feature = 'Wnt7a—Lrp6'
png(filename = paste(feature,'_CircuitPlot.png',sep=''),width = 5,height = 2.5,units = 'in',res=600)
TempFunc(feature = feature,
         max.edge.value = 0.25)
dev.off()

feature = 'Osm—Osmr'
png(filename = paste(feature,'_CircuitPlot.png',sep=''),width = 5,height = 2.5,units = 'in',res=600)
TempFunc(feature = feature,
         max.edge.value = 0.25)
dev.off()

# define new metadata slot for finding maximum edge value
merge <- merge(con.tri,con.quad)
merge$tempSlot <- paste(merge$Dataset2.Sending,merge$class.Joint)
temp <- AverageExpression(merge,group.by = 'tempSlot',slot = 'counts')
temp <- data.frame(temp$CellToCell)
temp$max <- apply(temp, 1, max, na.rm=TRUE)

# epithelium
for(i in 1:length(row.labels.epi)){
feature = row.labels.epi[i]
png(filename = paste(feature,'_CircuitPlot.png',sep=''),width = 5,height = 2.5,units = 'in',res=600)
TempFunc(feature = feature,
         max.edge.value = temp[row.labels.epi[i],]$max)
dev.off()
}

# Endothelium (borrows from other script, need to clean up)
for(i in 1:length(row.labels.end)){
feature = row.labels.end[i]
png(filename = paste(feature,'_CircuitPlot.png',sep=''),width = 5,height = 2.5,units = 'in',res=600)
TempFunc(feature = feature,
         max.edge.value = temp[row.labels.end[i],]$max)
dev.off()
}



# png(filename = paste(feature,'_FeaturePlot.png',sep=''),width = 5,height = 5,units = 'in',res=600)
# FeaturePlot(epi.rec,features= feature,label = T,max.cutoff = 1,min.cutoff = 0.25, order=T,repel = T)+NoAxes()
# dev.off()
```