---
title: "EpithelialReceivingWorking"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 24,fig.height = 22,warning = FALSE,message = FALSE)
```

## Introduction

This script analyzes epithelial niche signaling in the BEFM project

## Set WD
```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")
```

## Packages
```{r}
require(Seurat)
require(NICHES)
require(dplyr)
require(ggthemes)
require(ggplot2)
require(cowplot)
require(knitr)
```


# Local functions
```{r}
source('/Users/msbr/GitHub/NICHESMethods/R Scripts/ToolBuilding/NICHES_FeaturePlot.R')
```

#### Load saved NICHES data from Part 1

Note that no data filtration has been performed, per decisions in Part 2.

As a reminder, these data were made from a 5000 cell-per-sample downsampled object that is then fed into NICHES and the outputs saved. 

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
```

## Inspect
```{r}
global.connectomics
```


#### Subset out the just specific samples of interest, per Allie's request:

Here we limit all of our downstream analysis to information that came exclusively from the following 5 samples of interest, which will allow us to nicely isolate the effect of macrophages on mesenchymal to epithelial signaling. 

Note that this is a lossy step in this process, and that we might be losing useful information by doing this. But, we are doing this because it makes sense for this specific project - it allows us to asked focused questions within a data-context where we know that the answers will be reliable.

```{r}
samples.of.interest <- c('BEF12','BEF14','BEF15','BEFM1','BEFM2')
# we will use this set of samples for downstream calculations in this workflow
```

#### Define colors

This is a temporary color palette, and we should consider updating / polishing for publication.

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")

## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(global.connectomics$CellToCell$alra$SendingType)
```

#### Isolate Epithelial Receiving edges

Here, we isolated just the CellToCell signaling edges that land on Epithelium within this limited dataset. This is a newly discovered approach that seems to work well for this project. It allows us to ask questions like

"How does epithelial niche signaling change between conditions, who is responsible for the observed changes, and what features are associated with observed shifts?"


```{r}
temp <- global.connectomics$CellToCell$alra # we will use the imputed data, for now
Idents(temp) <- temp$class.Receiving
table(Idents(temp))
epi.rec <- subset(temp, idents = 'Epithelium')
Idents(epi.rec) <- epi.rec$orig.ident
table(Idents(epi.rec))
epi.rec <- subset(epi.rec,idents = c(samples.of.interest))
table(Idents(epi.rec))
```

## Clean

Since we are no longer prioritizing BCL5, we can clean these data to help us make prettier plots

```{r}
VlnPlot(epi.rec,'nFeature_CellToCell',group.by = 'orig.ident')
epi.rec <- subset(epi.rec,subset = nFeature_CellToCell>100)
VlnPlot(epi.rec,'nFeature_CellToCell',group.by = 'orig.ident')
```

Excellent. Let's take a global look at this edge category before moving further:

#### Epithelium Receiving - Embed and Cluster

```{r}
epi.rec <- ScaleData(epi.rec)
epi.rec <- FindVariableFeatures(epi.rec)
epi.rec <- RunPCA(epi.rec,npcs = 50)
ElbowPlot(epi.rec,ndims = 50)
epi.rec <- RunUMAP(epi.rec,dims = 1:20)
epi.rec <- FindNeighbors(epi.rec,dims = 1:20)
epi.rec <- FindClusters(epi.rec,resolution = 0.1)
```

## Epi Plot
```{r}
p1 <- DimPlot(epi.rec,group.by = 'class.Sending',shuffle = T)+ggtitle('Sending CellClass')
p2 <- DimPlot(epi.rec,group.by = 'class.Receiving',shuffle = T)+ggtitle('Receiving CellClass')
p3 <- DimPlot(epi.rec,group.by = 'Dataset2.Sending',shuffle = T)+ggtitle('Dataset2')
p4 <- DimPlot(epi.rec,group.by = 'SendingType',shuffle = T)+ggtitle('Sending CellType')
p5 <- DimPlot(epi.rec,group.by = 'ReceivingType',shuffle = T)+ggtitle('Receiving CellType')
p6 <- DimPlot(epi.rec,group.by = 'orig.ident',shuffle = T)+ggtitle('Sample')

# Define plot blocks
small.mult <- cowplot::plot_grid(p1,p2,p3,p4,p5,p6,ncol = 3)
sig.arch <- DimPlot(epi.rec,group.by = 'seurat_clusters',shuffle = T,label = T)+ggtitle('Signaling Archetypes')
split.arch <- DimPlot(epi.rec,split.by = 'orig.ident',group.by = 'seurat_clusters',shuffle = T,ncol = 2)+ggtitle('Signaling Archetypes (Split By Sample)')

# Assemble blocks in cowplot
top <- small.mult
bottom <- cowplot::plot_grid(sig.arch,split.arch,rel_widths = c(1.2,1))
plot <- cowplot::plot_grid(top,bottom,nrow = 2)
plot

png('Epithelial_Receiving.png',width=24,height = 22,units = 'in',res=300)
print(plot)
dev.off()
```

## Save for later
```{r}
save(epi.rec,file = 'epi.rec.Robj')
```
