---
title: "EpithelialReceivingWorking"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 20,fig.height = 10,warning = FALSE,message = FALSE)
```

#### Introduction

This script looks at every cell-to-cell signaling edge that lands on an epithelial cell.

#### Set Working Directory & Load Required Packages and Local Functions

Here we set our working directory and load our required packages for this analysis.

```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")

# Require packages
require(Seurat)
require(NICHES)
require(ggplot2)
require(cowplot)
require(dplyr)
require(knitr)

# Local functions
source('/Users/msbr/GitHub/NICHESMethods/R Scripts/ToolBuilding/NICHES_FeaturePlot.R')
```

#### Load saved NICHES data from Part 1

Note that no data filtration has been performed, per decisions in Part 2.

As a reminder, these data were made from a 5000 cell-per-sample downsampled object that is then fed into NICHES and the outputs saved. 

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
```

#### Subset out the just specific samples of interest, per Allie's request:

Here we limit all of our downstream analysis to information that came exclusively from the following 5 samples of interest, which will allow us to nicely isolate the effect of macrophages on mesenchymal to epithelial signaling. 

Note that this is a lossy step in this process, and that we might be losing useful information by doing this. But, we are doing this because it makes sense for this specific project - it allows us to asked focused questions within a data-context where we know that the answers will be reliable.

```{r}
samples.of.interest <- c('BEF12','BEF14','BEF15','BEFM1','BEFM2')
# we will use this set of samples for downstream calculations in this workflow
```

#### Define colors

This is a temporary color palette, and we should consider updating / polishing for publication.

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")

## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(global.connectomics$CellToCell$alra$SendingType)
```

#### Isolate Epithelial Receiving edges

Here, we isolated just the CellToCell signaling edges that land on Epithelium within this limited dataset. This is a newly discovered approach that seems to work well for this project. It allows us to ask questions like

"How does mesenchymal niche signaling change between conditions, who is responsible for the observed changes, and what features are associated with observed shifts?"


```{r}
temp <- global.connectomics$CellToCell$alra # we will use the imputed data, for now
Idents(temp) <- temp$class.Receiving
table(Idents(temp))
epi.rec <- subset(temp, idents = 'Epithelium')
Idents(epi.rec) <- epi.rec$orig.ident
table(Idents(epi.rec))
epi.rec <- subset(epi.rec,idents = c(samples.of.interest))
table(Idents(epi.rec))
```

Excellent. Let's take a global look at this edge category before moving further:

#### Epithelium Receiving

```{r}
epi.rec <- ScaleData(epi.rec)
epi.rec <- FindVariableFeatures(epi.rec)
epi.rec <- RunPCA(epi.rec,npcs = 100)
ElbowPlot(epi.rec,ndims = 100)
epi.rec <- RunUMAP(epi.rec,dims = 1:30)

p1 <- DimPlot(epi.rec,group.by = 'Dataset2.Sending',cols=color_pals$dataset_colors,raster=F,shuffle = T)
p2 <- DimPlot(epi.rec,group.by = 'orig.ident',cols = color_pals$sample_colors,raster=F,shuffle = T,label=T)
p3 <- DimPlot(epi.rec,group.by = 'class.Sending',cols = color_pals$class_colors,raster=F,shuffle = T)
p4 <- DimPlot(epi.rec,group.by = 'SendingType',cols = color_pals$type_colors,raster=F,shuffle = T)

plot_grid(p1,p2,p3,p4)

pdf(file='EpithelialReceiving.UMAPS.pdf',width=10,height=8)
print(plot_grid(p1,p2,p3,p4))
dev.off()
```

WELL this is just wild.

1. There is a distinct change in Epithelial autocrine signaling due to the addition of macrophages
2. There is a DIVERSIFICATION of Mesenchymal-to-Epithelial signals ---> WILD!!!!! I HAVE NEVER SEEN THIS BEFORE AND IT COMPLETELY FITS THE THEORY
3. Endothelia-to-Epitheliam signaling also diversifies a bit, and 'fills out' -- might need to look at this portion of the data within a narrowed latent space
4. Macrophage-to-Epithelial signaling is HIGHLY DISTINCT from inter-parenchymal signaling -- suggests that there are many intercellular signaling mechanisms that are highly distinct to macrophage-epithelial communication. 

Let's cluster these data and see if the clusters align with our intuition and support the above contentions:

```{r}
epi.rec <- FindNeighbors(epi.rec,dims = 1:30)
epi.rec <- FindClusters(epi.rec,resolution = 0.2)
DimPlot(epi.rec,label=T)

p5 <- DimPlot(epi.rec,label=T)
p6 <- DimPlot(epi.rec,group.by = 'ReceivingType',cols = color_pals$type_colors,raster=F,shuffle = T)
pdf(file='EpitheliumReceivingClustered.UMAPS.pdf',width=10,height=8)
print(plot_grid(p1,p3,p4,p2,p5,p6))
dev.off()
```

Great.

1. Cluster 2 captures the change in Epithelial-Epithelial signaling due to the addition of macrophages. [We should run differential expression between Clusters 2 and 1]
2. Cluster 5 captures the ADDITIONAL Mesenchymal-to-Epithelial signaling that exists due to the addition of macrophages [Might be BEFM1 specific?]
3. Cluster 6 is the macrophage-specific edges

Let's run a marker list across all archetypes to better understand feature-level changes that are driving the observed embedding and clustering patterns.

```{r}
mark <- FindAllMarkers(epi.rec,min.pct = 0.1,logfc.threshold = 0.1)
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
kable(mark %>% group_by(cluster) %>% top_n(20,ratio))
```

Cluster 5:
Wnt7a
Wnt7b
Npnt
Sct
Tnf
LAMININS !!!! We should look closely at this / read about this, I don't know what it might mean
Cgn?
Tgfa-Egfr
Mmp7???? why
Hbegf??? This is normally an ATI marker ligand...
Tgfb1/2?

Let's check out some feature plots to keep ourselves oriented on all of these / sanity check ourselves.

#### Load transcriptomic data so that we can make NICHES_FeaturePlots 

```{r}
 # Transcriptome object is called 'sub'
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/downsampled.engOnly.customEmbed.2023-12-14.Robj") # 'sub'
```

#### Set up data for NICHES_FeaturePlots

```{r}
# First we have to scale the connectivity data, just in case we haven't before
connectivity <- global.connectomics[['CellToCell']][['alra']]
connectivity <- ScaleData(connectivity)
```

We often want to split by Mono / Co / Tri / Quad. Currently we don't have a metadata slot that groups in this way. So, let's  create the appropriate metadata slot and organize the levels appropriately for plotting downstream. Note that we need to create parallel metadat slots in both the transcriptomic and connectomic objects.

```{r}
# Build a new metadata slot ('Dataset4') to split based on M/C/T/Q
Idents(sub) <- sub$Dataset2
Idents(connectivity) <- connectivity$Dataset2.Sending
table(Idents(sub))
table(Idents(connectivity))
sub <- RenameIdents(sub,
                       'Tri_E'='Tri',
                       'Tri_L'='Tri',
                       'Quad_E'='Quad',
                       'Quad_L'='Quad')
connectivity <- RenameIdents(connectivity,
                             'Tri_E'='Tri',
                             'Tri_L'='Tri',
                             'Quad_E'='Quad',
                             'Quad_L'='Quad')
sub$Dataset4 <- Idents(sub)
connectivity$Dataset4 <- Idents(connectivity)

sub$Dataset4 <- factor(sub$Dataset4,levels=c('Mono','Co','Tri','Quad'))
connectivity$Dataset4 <- factor(connectivity$Dataset4,levels=c('Mono','Co','Tri','Quad'))
```

Great. Now we are prepared to make a NICHES_FeaturePlot that splits by 'Dataset4'

```{r}
# Set up colors for legend plots
legend.palettes <- color_pals
# Make the names of legend.palettes correlates with the legends requested below
names(legend.palettes)[3] <- 'class'
names(legend.palettes)[1] <- 'Sample' 

# Split according to this new slot
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = 'Lamb3—Itgb1',
                       legends.to.plot = c('class','Sample'),
                       transpose = FALSE,
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```
Is this relevant? https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5677560/
This is a potentially interesting review: https://journals.sagepub.com/doi/abs/10.3181/00379727-214-44078

And we can experiment further:
```{r}
MOI <- 'Tgfa—Egfr'
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class','Sample'), 
                       transpose = FALSE,
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```

Interesting, what does this look like as a Violin Plot?

```{r}
p1 <- VlnPlot(connectivity,'Tgfa—Egfr',group.by = 'class.Joint',split.by = 'Dataset4',pt.size = 0,raster = FALSE)
p2 <- VlnPlot(connectivity,'Tgfa—Egfr',group.by = 'class.Joint',split.by = 'orig.ident',pt.size = 0,raster = FALSE)

plot_grid(p1,p2,nrow = 2)
```

Cool. Checks out and makes sense. So, adding macrophages promotes connectivity from epithelium to mesenchyme based on Tgfa-Egfr. We should probably research what this mechanism normally does / why it might be important.

Now, let's look at Cluster 5 markers:

Fn1—Itgb6 - integrin beta6 is one of the top defining markers of pulmonary fibrosis. how does fibronectin interact with it / why would it be relevant?

Maybe relevant? https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2267338/
https://www.atsjournals.org/doi/full/10.1164/rccm.200708-1291OC

Thbs1 - thrombospondin 1, very close to what themis studies, extracellular regulator of fibrosis and wound healing...

Il6—F3 -- -what does this do biologicaly?

Fgf2—Fgfr3 WOW COOL

Wnt16—Fzd6 IS this real? need to check non-imputed data before we run with a random unknown WNT
Reln—Itga3 might be related to lymphatic or pericytes differentiation?
Mdk—Tspan1 could be relevant
Mmp9—Ephb2 interesting or not?
Fgf7—Fgfr3 GEEZ louise
Ngf—Kidins220 I don't know what this does but it seems interesting
Fbn1—Itgb6 yikes this is strongly upregulated
Tnc—Egfr oh baby cool
Tfpi—F3 what does this mechanism do ?
Tgfb1 and assorted receptors strongly up in Quad

Hbegf—Egfr mmmhmm this is here too, (in mes-epi as well as epi-epi)
Inhba—Acvr2b this is an extracellular modulator of TGFB signaling, worth reading about a bit i think
Wnt5a and assorted receptors -- amazing -- pro-alveologenesis cue and pro-alveolar epithelial differentiation cue, confirmed to be critical for the resolution of fibrosis in mice: Konigschoof's work
Bmp2—Bmpr1a -- oo i bet this is cool and real

Let's check a few:


```{r}
MOI <- 'Wnt5a—Fzd6'
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class','Sample'),
                       transpose = FALSE, 
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
MOI <- 'Wnt5a—Lrp5'
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class','Sample'),
                       transpose = FALSE, 
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
MOI <- 'Wnt5a—Ryk'
NICHES_FeaturePlot(transcriptome.object = sub,
                       connectome.object = connectivity,
                       mechanism.of.interest = MOI,
                       legends.to.plot = c('class','Sample'),
                       transpose = FALSE, 
                       split.by = 'Dataset4',
                       legend.palettes = legend.palettes,
                       min.connectivity.thresh = 0.1)
```

Fascinating. Look at where the yellow lines are concentrated. This is a major pro-alveologenesis mechanism, known in mouse, that we have directly replicated in our engineered system. Allie. Amazing.

#### Conclusions:

This is a powerful way to look at the data.
