---
title: "MesEpiSubset_2023-11-11"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 20,fig.height = 10,warning = FALSE,message = FALSE)
```

This script is a quick-and-dirty workflow to look specifically at mesenchymal to epithelial signaling in the BEFM project data.

EDITED TO JUST LOOK AT BEF12, BEF14, BEF15, BEFM1 and BEFM2 per Allie's request 2023-11-08

We will do this three specific ways: we will look at Mesenchymal influence, the Epithelial microenvironment, and Mesencymal-Epithelial signaling. These three lenses should allow us to clearly understand what is going on.

If we like what it produces, it may be a useful standardized framework to look at the full dataset, later. This workflow is experimental for now.

#### Setup Packages

```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")

# Require packages
require(Seurat)
require(NICHES)
require(ggplot2)
require(cowplot)
require(dplyr)
require(knitr)
```

#### Load NICHES data from Part 1

These data have not been filtered, per Part 2 decisions. It seems to be OK.

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/global.connectomics.2023-11-11.Robj")
```


#### subset out the just specific samples of interest, per Allie's request:
```{r}
samples.of.interest <- c('BEF12','BEF14','BEF15','BEFM1','BEFM2')
# we will use this downstream
```

#### Define colors

This is a temporary color palette, and should be updated for publication.

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")

## For now:
color_pals$type_colors <- c('#A40606','#9CFFFA','#B0DB43','#9C528B','#2F6690',
                                '#946846','#F1C40F','green','#0F0326','#E65F5C','#14591D','#726DA8',
                                'yellow','purple','blue','red','orange','darkgrey','magenta')
names(color_pals$type_colors) <- unique(global.connectomics$CellToCell$alra$SendingType)
```

#### Isolate Mesenchymal Influence

This is just the Mesenchymal Cell-To-System data, which incorporate information about both cell-signaling character and cell population presence / absence / representation.

```{r}
temp <- global.connectomics$CellToSystem$alra # let's just look at the imputed data, for now, for ease
Idents(temp) <- temp$class
table(Idents(temp))
mes.inf <- subset(temp, idents = 'Mesenchyme')
Idents(mes.inf) <- mes.inf$orig.ident
table(Idents(mes.inf))
mes.inf <- subset(mes.inf,idents = c(samples.of.interest))
table(Idents(mes.inf))
```

#### Isolate Epithelial Microenvironment

This is just the Epithelial System-To-Cell data, which incorporates information about all signaling landing on Epithelium, regardless of source. It changes, like above, both with changing cellular phenotype and with altered cell population representation.

```{r}
temp <- global.connectomics$SystemToCell$alra
Idents(temp) <- temp$class
table(Idents(temp))
epi.mic <- subset(temp, idents = 'Epithelium')
Idents(epi.mic) <- epi.mic$orig.ident
table(Idents(epi.mic))
epi.mic <- subset(epi.mic,idents = c(samples.of.interest))
table(Idents(epi.mic))
```

#### Isolate Mesenchymal to Epithelial Signaling

This just looks at the Mesenchymal-To-Epithelium Cell-To-Cell communication. It predominantly captures signal due to altered cellular character. It does not appropriately capture altered cell-signaling cues due to altered cell population representation, but it MAY capture signal present in this cross DUE to the addition of unseen cells outside of this cross (i.e., it could change with/without the presence of macrophages, even though there will be no data FROM macrophages in this object.)

```{r}
temp <- global.connectomics$CellToCell$alra
Idents(temp) <- temp$class.Joint
table(Idents(temp))
mes.epi <- subset(temp,idents = 'Mesenchyme - Epithelium')
Idents(mes.epi) <- mes.epi$orig.ident
table(Idents(mes.epi))
mes.epi <- subset(mes.epi,idents = c(samples.of.interest))
table(Idents(mes.epi))
```

Ok, now we have isolated our data of interest. Next, let's visualize each data set:

#### Mesenchymal Influence

```{r}
mes.inf <- ScaleData(mes.inf)
mes.inf <- FindVariableFeatures(mes.inf)
mes.inf <- RunPCA(mes.inf,npcs = 100)
ElbowPlot(mes.inf,ndims = 100)
mes.inf <- RunUMAP(mes.inf,dims = 1:25)

p1 <- DimPlot(mes.inf,group.by = 'Dataset2',cols=color_pals$dataset_colors,raster=F,shuffle = T)
p2 <- DimPlot(mes.inf,group.by = 'orig.ident',cols = color_pals$sample_colors,raster=F,shuffle = T,label=T)
p3 <- DimPlot(mes.inf,group.by = 'class',cols = color_pals$class_colors,raster=F,shuffle = T)
p4 <- DimPlot(mes.inf,group.by = 'SendingType',cols = color_pals$type_colors,raster=F,shuffle = T)

pdf(file='MesenchymeInfluence_Subset.UMAPS.pdf',width=10,height=8)
print(plot_grid(p1,p2,p3,p4))
dev.off()

plot_grid(p1,p2,p3,p4)
```

I think this look pretty good. Significant tissue-tissue separation, which may be a consequence of real biology or may be batch, we do not know yet. We could try integrating these later if we would like, not sure if it would help us or obscure interesting patterns. Let's avoid for the moment, but can revisit this decision.

#### Epithelial Microenvironment

```{r}
epi.mic <- ScaleData(epi.mic)
epi.mic <- FindVariableFeatures(epi.mic)
epi.mic <- RunPCA(epi.mic,npcs = 100)
ElbowPlot(epi.mic,ndims = 100)
epi.mic <- RunUMAP(epi.mic,dims = 1:25)

p1 <- DimPlot(epi.mic,group.by = 'Dataset2',cols=color_pals$dataset_colors,raster=F,shuffle = T)
p2 <- DimPlot(epi.mic,group.by = 'orig.ident',cols = color_pals$sample_colors,raster=F,shuffle = T,label=T)
p3 <- DimPlot(epi.mic,group.by = 'class',cols = color_pals$class_colors,raster=F,shuffle = T)
p4 <- DimPlot(epi.mic,group.by = 'ReceivingType',cols = color_pals$type_colors,raster=F,shuffle = T)

pdf(file='EpithelialMicroenviornment_Subset.UMAPS.pdf',width=10,height=8)
print(plot_grid(p1,p2,p3,p4))
dev.off()

plot_grid(p1,p2,p3,p4)
```

This is actually pretty intresting and I hven't noticed these patterns before. The tissues group more together here. The similarity between Mono + Tri_E vs. Co + Tri_L + Quad is interesting and I suspect relevant. 

EDITS 11/9/2023, post-subset: BEFM1 is clearly different from the rest. Also this is too few data points. so I need to go back to Part 1 and change my downsampling approaches.

#### Mesenchymal-Epithelial Communication

```{r}
mes.epi <- ScaleData(mes.epi)
mes.epi <- FindVariableFeatures(mes.epi)
mes.epi <- RunPCA(mes.epi,npcs = 100)
ElbowPlot(mes.epi,ndims = 100)
mes.epi <- RunUMAP(mes.epi,dims = 1:25)

p1 <- DimPlot(mes.epi,group.by = 'Dataset2.Sending',cols=color_pals$dataset_colors,raster=F,shuffle = T)
p2 <- DimPlot(mes.epi,group.by = 'orig.ident',cols = color_pals$sample_colors,raster=F,shuffle = T)
p3 <- DimPlot(mes.epi,group.by = 'class.Sending',cols = color_pals$class_colors,raster=F,shuffle = T)
p4 <- DimPlot(mes.epi,group.by = 'class.Receiving',cols = color_pals$class_colors,raster=F,shuffle = T)
p5 <- DimPlot(mes.epi,group.by = 'SendingType',cols = color_pals$type_colors,raster=F,shuffle = T)
p6 <- DimPlot(mes.epi,group.by = 'ReceivingType',cols = color_pals$type_colors,raster=F,shuffle = T)

pdf(file='MesenchymalToEpitheliumCommunication_Subset.UMAPS.pdf',width=20,height=10)
print(plot_grid(p1,p2,p3,p4,p5,p6))
dev.off()

plot_grid(p1,p2,p3,p4,p5,p6)
```

This is very cool! Looks like there is a clear effect on Mesenchymal-Epithelial communication due to the addition of Macrophages. Let's dig into this specific signal a bit by running a statistical test comparing these two categories:

EDITS 11/9/2023, post-subset: Couple things here. 1. There is a clear association between alveolar fibroblasts with Quad culture, that is already known I think from your analysis but it is cool to see here. 2. Seems like aberrant fibroblasts go down with the addition of macrophages, which again is known from your analysis but i also think is cool. 3. Is quad culture associated with transitional epithelium? Overarching: definitely different Mes-Epi signaling associated with quad culture, even in this limited subset system of tissues.

#### Compare MesEpi with and without macrophages:

First, we want a new meta data category that annotates Tri culture vs. Quad culture. we do this as follows: 

```{r}
mes.epi$Dataset4 <- NA
Idents(mes.epi) <- mes.epi$Dataset2.Sending 
mes.epi <- RenameIdents(mes.epi,
                        'Tri_E'='Tri',
                        'Tri_L'='Tri',
                        'Quad_E'='Quad',
                        'Quad_L'='Quad')
mes.epi$Dataset4 <- Idents(mes.epi)
table(mes.epi$Dataset4)
```

Now, we can run a marker test comparing these two categories specifically:

```{r}
mark.mes.epi <- FindAllMarkers(mes.epi,only.pos = T,min.pct = 0.1,logfc.threshold = 0.01)
mark.mes.epi$ratio <- mark.mes.epi$pct.1/mark.mes.epi$pct.2
mark.mes.epi$power <- mark.mes.epi$ratio*mark.mes.epi$avg_log2FC
kable(mark.mes.epi %>% group_by(cluster) %>% top_n(50,ratio) %>% arrange(desc(ratio)))
```

Not bad! Some very interesting signals in there. Let's remove signaling mediated by integrins, which are less interesting for this project.

```{r}
MOI <- mark.mes.epi[-grep('Itg',mark.mes.epi$gene),]
MOI <- MOI %>% group_by(cluster) %>% top_n(50,ratio) %>% arrange(desc(ratio))
kable(MOI)
```

So there is definitely a huge activation of WNT family signaling here. Wnt7a, Wnt7b, via Lrp5/6. Rspo1-Lgr4 (I know the ligand well, but not the receptor!) Also mild upregulation of Fgf1/2/7 signaling via Fgfr3. Also a big upregulation of Hbegf-Egfr. The tri-culture elevation of Dlk1-Notch1 is interesting, not sure what it means. Nrg1-Erbb3 also upregulated in Quad, and Erbb3 is an powerful epithelial stem-cell / growth activator. The Csf2-Csf2ra upregulated in Quad might be relevant to macrophage biology. Efna4—Epha7 also upregulated, it is a spatial guidance molecule implicated in tissue morphogenesis, as is Sema4b—Dcbld2, which is acting through a receptor I have never seen before. Il11—Il6st also upregulated. Immune recruitment / activation? Not sure what this mechanism does, would need to read.

#### Let's make some feature plots of Mes-Epi signals that are upregulated in Quad culture as compared to Tri, like this:

```{r warning=FALSE}
quick.function <- function(mech){
p1 <- FeaturePlot(mes.epi,mech)
p2 <- VlnPlot(mes.epi,mech)
p3 <- VlnPlot(mes.epi,mech,group.by = 'orig.ident')
plot_grid(p1,p2,p3,nrow = 1)
}
mech <- 'Fgf7—Fgfr3'
quick.function(mech)

mech <- 'Wnt7a—Lrp6'
quick.function(mech)

mech <- 'Wnt7b—Lrp5'
quick.function(mech)

mech <- 'Csf2—Csf2ra'
quick.function(mech)

mech <- 'Fgf2—Fgfr3'
quick.function(mech)

mech <- 'Fgf2—Fgfr2'
quick.function(mech)

mech <- 'Hbegf—Egfr'
quick.function(mech)

mech <- 'Il11—Il6st'
quick.function(mech)

mech <- 'Rspo1—Lgr4'
quick.function(mech)

```


EDITS 11/9/2023, post-subset: I have added a third plot here that looks at the data split by Tissue, so that we can see if any of the effects are sample-specific. So, while these mechanisms are definitely still interesting, I am seeing the same thing you have mentioned, which is that BEFM1 is like "much better" than other quad cultures. Weird. But I guess also makes sense? We already kind of knew this. Not a huge issue, the tissue looks the best too...

#### Next, let's look at some Tri culture Mes-Epi markers:

Not sure what Dlk1 is doing or if it is relevant, it is a Notch ligand and it is definitely lowered in BEFM1: 

```{r warning=FALSE}
mech <- 'Dlk1—Notch1'
quick.function(mech)
```

Thbs2 is Themis' pro-fibrotic, anti-angiogenic matrix molecule that he knocks out, wild that it is here so clearly and is so reduced in BEFM1 and the quad cultures more generally

```{r}
mech <- 'Thbs2—Cd47'
quick.function(mech)
```

Sfrp1—Fzd6 is an extracellular binder of WNT ligands that sequesters them outside of cells and rpevents them from binding, it is net anti-WNT. It is DOWNREGULATED in quad culture, suggesting that macrophage addition will activate this pathway and promote epithelial WNT signal transduction

```{r}
mech <-'Sfrp1—Fzd6'
quick.function(mech)
```

Cthrc1—Fzd6 is a major matrix molecule implicated in fibrosis (ligand): https://www.nature.com/articles/s41467-020-15647-5 and the receptor is a critical cue for alveolar regeneration by epithelium: https://www.cell.com/cell/pdf/S0092-8674(23)00541-X.pdf

```{r}
mech <-'Cthrc1—Fzd6'
quick.function(mech)
```

#### My summary, in a nutshell:

Tri-culture has major upregualtion of fibrotic pathways, genes, and matrix molecules. Adding macrophages basically removes this signal.

Adding macrophages in Quad-culture promotes significant upregulation of WNT-signaling and FGF-family signaling. We also see the expression of prominent sptial-guidance molecules crucial to local tissue organziation that we know from normal alveolar homeostasis.

#### Comments, post-subsetting down to just samples of interest:

1. This will look better once we go back to Part 1 and decrease our downsampling operation. We have about 5 times as many total possible  measurements to look at for every cateogry, I have deliberately limited our analyses to this point to about 1/5 of the data to speed computation and allow us to do this exploratory analysis together before going back and running the whole dataset. I may do that next, because this is now a very small chunk of the data we are considering and I would like it to be more fleshed out and comprehensive.


