---
title: "MesRecFeatures_v2"
author: "MSBR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 24,fig.height = 22,warning = FALSE,message = FALSE)
```

## Introduction

This script identifies specific features of interest, landing on MESENCHYME, in the BEFM project.

## Set WD
```{r message=F}
# Set WD
setwd("/Users/msbr/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics")
```

## Packages
```{r}
require(Seurat)
require(NICHES)
require(dplyr)
require(ggthemes)
require(ggplot2)
require(cowplot)
require(knitr)
```

## Color palette

```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Allie's Object Filtration/Final Objects/all.color_palettes_2.R")
color_pals
```

## Load Mes Rec from save
```{r}
load("~/Library/CloudStorage/GoogleDrive-michasam.raredon@yale.edu/.shortcut-targets-by-id/1VLqBlyzO-Qad5O2kbwXkBRh_1cQho39t/Engineered Lung Paper/Global_Connectomics/mes.rec.Robj")
```

## Inspect
```{r}
mes.rec
```

## Update to v5
```{R}
mes.rec <- UpdateSeuratObject(mes.rec)
mes.rec
```
## Dimplot for review
```{r}
p0 <- DimPlot(
  mes.rec,
  reduction = "umap",
  group.by = c('class.Sending',"Dataset2.Sending", "orig.ident", "seurat_clusters"),
  combine = FALSE, label.size = 8,shuffle = T,label=T)
p0 <- cowplot::plot_grid(plotlist = p0,nrow = 1)
p0
```
## Define archetypes for analysis
```{r}
mes.rec$Archetype <- mes.rec$seurat_clusters
Idents(mes.rec) <- mes.rec$Archetype
```

## Dimplot for review
```{r}
p0 <- DimPlot(
  mes.rec,
  reduction = "umap",
  group.by = c('Archetype','class.Sending',"Dataset2.Sending", "orig.ident", "pca_clusters"),
  combine = FALSE, label.size = 8,shuffle = T,label=T)
p0 <- cowplot::plot_grid(plotlist = p0,nrow = 1)
p0
```

## Label based on metdata

```{r}
Idents(mes.rec) <- mes.rec$seurat_clusters
table(Idents(mes.rec))
mes.rec <- RenameIdents(mes.rec,
                        '0'='Mes-Auto BEF12',
                        '1'='Baseline Epi-Mes',
                        '2'='Mes-Auto BEF14',
                        '3'='Mes-Auto BEF15',
                        '4'='Mes-Auto BEFM2',
                        '5'='Baseline Endo-Mes',
                        '6'='Epi-Mes BEFM1',
                        '7'='Mes-Auto BEFM1',
                        '8'='Immune-Mes')
table(Idents(mes.rec))
mes.rec$Archetype <- Idents(mes.rec)
```


# Differential expression across archetype

```{r}
mark <- FindAllMarkers(mes.rec,min.pct = 0.1,logfc.threshold = 0.1)
mark$ratio <- mark$pct.1/mark$pct.2
mark$power <- mark$ratio*mark$avg_log2FC
kable(mark %>% group_by(cluster) %>% top_n(20,ratio))
```

## Cluster 0 (Mes-Auto BEF12)
```{r}
goi.0 <-  c()
FeaturePlot(mes.rec,
            features = goi.0,order = T,pt.size=1,label=T,label.size = 8)

```

## Cluster 1 (Baseline Epi-Mes)
```{r}
goi.1 <-  c()
FeaturePlot(mes.rec,
            features = goi.1,order = T,pt.size=1,label=T,label.size = 8)

```

## Cluster 2 (Mes-Auto BEF14)
```{r}
goi.2 <-  c()
FeaturePlot(mes.rec,
            features = goi.2,order = T,pt.size=1,label=T,label.size = 8)

```

## Cluster 3 (Mes-Auto BEF15)
```{r}
goi.3 <- c()
FeaturePlot(mes.rec,
            features = goi.3,
            order = T,pt.size = 1,label=T,label.size = 8)
```

## Cluster 4 (Mes-Auto BEFM2)
```{r}
goi.4 <- c()
  
FeaturePlot(mes.rec,
            features = goi.4,
            order = T,pt.size = 1,label=T,label.size = 8)
```

## Cluster 5 (Baseline Endo-Mes)
```{r}
goi.5 <- c('Dll4—Notch2','Vip—Npr3','Tnfsf4—Traf2','Tnfsf10—Tnfrsf11b','Nxph1—Nrxn3','Angpt2—Tie1','Mfng—Notch2',
'Pdgfb—Pdgfrb','Pdgfb—Pdgfra','Pdgfd—Pdgfrb','Dhh—Ptch1','Dhh—Cdon','Dhh—Boc','Bmp4—Bmpr1a','Bmp4—Bmpr2','Sema6d—Plxna1')
  
FeaturePlot(mes.rec,
            features = goi.5,
            order = T)
```

## Cluster 6 (Epi-Mes BEFM1)
```{r}
goi.6 <- c('Wnt7a—Lrp6','Fgf9—Fgfr2','Tgfb1—Itgb6','Tnf—Ltbr','Bmp7—Bmpr1b','Npnt—Itgb1','Sct—Sctr','Cdh1—Igf1r','Mmp7—Cd44')
  
FeaturePlot(mes.rec,
            features = goi.6,
            order = T)
```

## Cluster 7 (Mes-Auto BEFM1)
```{r}
goi.7 <- c('Fgf1—Fgfr2','Fbn1—Itgb6','Mdk—Tspan1','Fgf2—Fgfr2','Tgfb1—Itgb6','Fn1—Itgb6')
  
FeaturePlot(mes.rec,
            features = goi.7,
            order = T)
```

## Cluster 8 (Immune-Mes)
```{r}
goi.8 <- c()
  
FeaturePlot(mes.rec,
            features = goi.8,
            order = T,pt.size = 1,label=T,label.size = 8)
```

## Heatmap

```{r}
goi.0 <-  c('Dlk1—Notch1','Bmp5—Bmpr1b','Rspo3—Lrp6',
                         'Angpt1—Itgb1','Bmp5—Acvr2a','Thbs2—Itgb1',
                         'Bmp5—Bmpr1a','Thbs2—Cd47','Ptn—Sdc1','Mst1—Mst1r','Fgf7—Fgfr1','Igf2—Igf2r','Bmp5—Bmpr1b','Ptn—Sdc1','Fgf10—Fgfr2','Dkk2—Lrp6','Bmp5—Bmpr2')

goi.1 <-  c('Efna3—Epha1','Efna3—Epha2','Efna3—Epha4','Ntf4—Ngfr','Il10—Il10rb','Fgf9—Fgfr1','Dlk2—Notch1','Wnt3a—Fzd6','Wnt3a—Lrp6','Wnt3a—Ryk','Fgf9—Fgfr2','Bmp7—Eng','Bmp7—Bmpr1b','Wnt7b—Fzd1','Bmp7—Bmpr2','Areg—Erbb3','Ereg—Erbb3','Areg—Egfr')

goi.2 <-  c('Jag1—Notch4','Dll1—Notch3','Bmp3—Bmpr1b','Dll1—Notch4','Gas6—Axl','Tnfsf12—Tnfrsf25','Jag1—Notch3','Bmp6—Bmpr2','Bmp3—Bmpr2','Vegfc—Nrp2','Vip—Sctr','Dll1—Notch2','Dll1—Notch1','Tnfsf4—Tnfrsf4','Dll4—Notch1','Dll4—Notch2','Pdgfb—Pdgfrb','Angpt2—Tie1','Dll4—Notch3','Dll4—Notch4','Dhh—Ptch1','Sema6d—Plxna1')

goi.3 <- c('Tnf—Tnfrsf1b','Sct—Sctr','Wnt7a—Lrp6',
                         'Npnt—Itgb1','Ereg—Egfr','Wnt7b—Lrp5',
                         'Hbegf—Egfr','Lamb3—Itga3','Npnt—Itgb1','Tnf—Tnfrsf21','Tnf—Tnfrsf21','Tnf—Ltbr','Lamb3—Itga2','Mmp7—Cd44','Lamb3—Itga2','Btc—Egfr')

goi.4 <- c('Fgf1—Fgfr3','Fgf1—Egfr','Fgf1—Fgfr2','Fgf2—Fgfr3','Fgf12—Fgfr3','Lama1—Itga2',
                         'Csf2—Csf2ra','Ngf—Maged1','Wnt16—Fzd6',
                         'Fgf7—Fgfr3','Fn1—Itgb6','Vegfa—Ephb2',
                         'Wnt5a—Lrp5','Wnt5a—Ryk','Col4a1—Itga2','Tnc—Egfr')
  
goi.5 <- c('Amh—Egfr','Hgf—Met','Igf1—Igf1r','Cxcl2—Cxcr2','Il1b—Adrb2','Tnfsf15—Tnfrsf25','Ebi3—Il6st','Il18—Il18r1','Osm—Osmr','Il1b—Il1r1','Camp—Igf1r','Tnfsf13—Tnfrsf1a','Nrg4—Egfr','Nrg4—Erbb2','Efna2—Epha1')
  
goi <- unique(c(goi.0,goi.1,goi.2,goi.3,goi.4,goi.5))

row.labels <- c('Bmp5—Bmpr1b','Rspo3—Lrp6','Fgf10—Fgfr2',
                'Il10—Il10rb','Wnt3a—Fzd6','Wnt7b—Fzd1',
                'Dll1—Notch2','Vip—Sctr','Dhh—Ptch1',
                'Sct—Sctr','Wnt7a—Lrp6','Wnt7b—Lrp5',"Lamb3—Itga3",'Tnf—Ltbr',
                'Fgf7—Fgfr3','Fn1—Itgb6','Col4a1—Itga2','Wnt5a—Ryk',
                'Osm—Osmr','Il1b—Il1r1','Il18—Il18r1','Nrg4—Egfr')

png('EpiRec.Heatmap.6.png',width=8,height=8,units='in',res=300)
CustomHeatmap(mes.rec,
              data.type = 'CellToCell',
                 primary = 'Archetype' ,
                 secondary = 'class.Sending' ,
                 tertiary = 'Dataset2.Sending' ,
                 quarternary = 'orig.ident' ,
                 primary.cols = NULL ,
                 secondary.cols = color_pals$class_colors,
                 tertiary.cols = color_pals$dataset_colors[5:6],
                 quarternary.cols = color_pals$sample_colors,
                 features = goi,
                 labels = c('Signaling Archetype','Sending Cell Class','Condition','Sample'),
                selected.row.anotations=row.labels,
              selected.label.size = 10)
dev.off()
```
